% paper/checklist.tex
% NeurIPS 2026 Datasets and Benchmarks Track --- Paper Checklist.
% Each item is followed by [Yes]/[No]/[NA] and a short justification.

\section*{NeurIPS Paper Checklist}

\begin{enumerate}

\item \textbf{Claims.}
\emph{Do the main claims made in the abstract and introduction
accurately reflect the paper's contributions and scope?}\\
\textbf{Answer:} [Yes] \\
\textbf{Justification:} The abstract and introduction describe a
multi-task benchmarking methodology (schema extraction, numerical
grounding, evidence attribution) instantiated on a nanobubble case
study, three metrics that operationalise three failure modes, a
reproducible release with two annotation tiers (gold-hard,
gold-lite), and three baselines (B1 regex, B2 BioBERT-SQuAD-v2,
B3 \texttt{Qwen2.5-7B-Instruct}). All numerical claims in the abstract
are supported by the headline-results section
(\S\ref{sec:headline}) and are scored against the gold-hard tier. We
explicitly declare the abstract-only scope and the v1.0 versus v1.1
boundary in \S\ref{sec:limitations}.

\item \textbf{Limitations.}
\emph{Does the paper discuss the limitations of the work performed
by the authors?}\\
\textbf{Answer:} [Yes] \\
\textbf{Justification:} \S\ref{sec:limitations} discusses
single-domain instantiation, abstract-only scope, two-tier annotation
(gold-hard tier single-annotator-validated under a pre-specified
dual-annotator protocol with full dual annotation planned for v1.1;
gold-lite tier single-annotator throughout), baseline scope (no
proprietary frontier models), and the licensing posture. The
annotation-status paragraph in \S\ref{sec:validity} states honestly
that v1.0 ships single-annotator labels under the protocol.

\item \textbf{Theory assumptions and proofs.}
\emph{For each theoretical result, does the paper provide the full
set of assumptions and a complete (and correct) proof?}\\
\textbf{Answer:} [NA] \\
\textbf{Justification:} The paper introduces no formal theoretical
results; the metrics in \S\ref{sec:experimental_setup} are operational
definitions, not theorems.

\item \textbf{Experimental result reproducibility.}
\emph{Does the paper fully disclose all the information needed to
reproduce the main experimental results?}\\
\textbf{Answer:} [Yes] \\
\textbf{Justification:} Section~\ref{sec:experimental_setup} specifies the model
checkpoints, decoding parameters, and confidence thresholds for B1,
B2, and B3. Appendix~\ref{app:pipeline} documents the harvest query
families, deduplication order, tier rules, and benchmark candidate
sampling parameters. The dev/test split is constructed
deterministically (seed=42); the split file, slice-membership map,
and leakage-audit report ship with the release.

\item \textbf{Open access to data and code.}
\emph{Does the paper provide open access to the data and code, with
sufficient instructions to faithfully reproduce the main experimental
results?}\\
\textbf{Answer:} [Yes] \\
\textbf{Justification:} The release at the public URLs in
\S\ref{sec:release} (GitHub for code, HuggingFace for data) currently
ships the Python package (\texttt{nanobubbleval}), four numbered
pipeline scripts that reproduce every artifact end-to-end, the
40-record gold-hard tier with first-annotator labels, the three v1.0
baseline prediction files, a Croissant-1.1 metadata file, the schema
specification, and the slice and leakage reports. The remaining
dataset artifacts (52{,}519-record warehouse manifest, 8{,}006-record
high-precision core, 460-record gold-lite extension tier, three task
views, and the deterministic \texttt{splits.json}) are scheduled for
the full release before the camera-ready version. The package ships
with a unit-test suite (\texttt{tests/}) covering the schema,
normaliser, evaluator, reconciler, splits builder, baselines, and
harvest framework.

\item \textbf{Experimental setting / details.}
\emph{Does the paper specify all the training and test details
(e.g., data splits, hyperparameters, how they were chosen, type of
optimizer, etc.) necessary to understand the results?}\\
\textbf{Answer:} [Yes] \\
\textbf{Justification:} Section~\ref{sec:experimental_setup} specifies the model
identifier, decoding strategy (greedy), and confidence threshold
($0.20$) for each baseline. Threshold tuning is performed on the
gold-lite dev split; test reporting uses the frozen value.
Section~\ref{sec:experimental_setup} defines all metrics, including the
relative tolerance ($\tau=0.05$) for tolerance-bounded numerical
match. No baseline is fine-tuned in v1.0; this is a deliberate scope
decision documented in \S\ref{sec:experimental_setup}.

\item \textbf{Experiment statistical significance.}
\emph{Does the paper report error bars suitably and correctly defined
or other appropriate information about the statistical significance
of the experiments?}\\
\textbf{Answer:} [Yes] \\
\textbf{Justification:} Section~\ref{sec:slices} and
Appendix~\ref{app:errors} report bootstrap 95\% confidence intervals
on the headline macro Acal-F1 and the n=8 review-paper slice. We
report point estimates with bootstrap CIs rather than significance
tests because the gold-hard tier is a fixed evaluation set rather
than a random sample from a larger population.

\item \textbf{Experiments compute resources.}
\emph{For each experiment, does the paper provide sufficient
information on the computer resources (type of compute workers,
memory, time of execution) needed to reproduce the experiments?}\\
\textbf{Answer:} [Yes] \\
\textbf{Justification:} Section~\ref{sec:experimental_setup} states that B3
inference runs on a single 16~GB GPU with greedy decoding (40-record
inference completes in approximately 7~minutes wall-clock on an
NVIDIA TITAN RTX). B2 inference completes in under 15~seconds on the
same hardware. B1 has no model dependency.

\item \textbf{Code of ethics.}
\emph{Does the research conducted in the paper conform, in every
respect, with the NeurIPS Code of Ethics?}\\
\textbf{Answer:} [Yes] \\
\textbf{Justification:} The benchmark contains no human subjects, no
identifiable medical records, and no patient-level information. All
warehouse records are sourced from public scholarly APIs under their
respective terms; the release preserves DOIs and PMIDs and
redistributes only derived structured curation. Annotation was
performed by adult professional annotators under informed agreement.

\item \textbf{Broader impacts.}
\emph{Does the paper discuss both potential positive societal impacts
and negative societal impacts of the work performed?}\\
\textbf{Answer:} [Yes] \\
\textbf{Justification:} \S\ref{sec:impact} discusses the intended
positive impact (exposing three jointly stressed failure modes in
scientific information extraction so that downstream pipelines such
as literature review, database construction, and meta-analysis do
not act on unsupported emissions) and two principal risks
(extrapolation beyond the calibrated nanobubble domain;
English-language source bias). The mitigations baked into the design
(open-source schema and evaluator, documented tier-curation and
slice-tagging, semantic versioning) are also described there.

\item \textbf{Safeguards.}
\emph{Does the paper describe safeguards that have been put in place
for responsible release of data or models with a high risk for misuse
(e.g., pretrained language models, image generators, or scraped
datasets)?}\\
\textbf{Answer:} [NA] \\
\textbf{Justification:} The release contains structured metadata and
abstract-derived annotations only; it does not redistribute any model
weights, full-text papers, or scraped commercial content. Risk of
misuse is comparable to that of existing biomedical-extraction
benchmarks.

\item \textbf{Licenses for existing assets.}
\emph{Are the creators or original owners of assets (e.g., code,
data, models) used in the paper properly credited and are the license
and terms of use explicitly mentioned and properly respected?}\\
\textbf{Answer:} [Yes] \\
\textbf{Justification:} PubMed/Entrez, Europe~PMC, OpenAlex, CrossRef,
Semantic~Scholar, and ClinicalTrials.gov are cited as data sources;
their respective terms of use permit redistribution of bibliographic
metadata and abstracts. The B2 model
(\texttt{ktrapeznikov/biobert\_v1.1\_pubmed\_squad\_v2}) is released
on HuggingFace under its model card terms; B3
(\texttt{Qwen/Qwen2.5-7B-Instruct}) is released under the Qwen
license. Model identifiers are reported verbatim in
\S\ref{sec:experimental_setup}.

\item \textbf{New assets.}
\emph{Are new assets introduced in the paper well documented and is
the documentation provided alongside the assets?}\\
\textbf{Answer:} [Yes] \\
\textbf{Justification:} The release ships a dataset card
(Appendix~\ref{app:datasheet}), the Croissant-1.1 metadata file with
NeurIPS-required RAI fields, a schema specification
(Appendix~\ref{app:schema}), the annotation guideline used for the
IAA pass (Appendix~\ref{app:guideline}), and a release card listing
canonical file paths (Appendix~\ref{app:release_card}).

\item \textbf{Crowdsourcing and research with human subjects.}
\emph{For crowdsourcing experiments and research with human subjects,
does the paper include the full text of instructions given to
participants and screenshots, if applicable, as well as details
about compensation (if any)?}\\
\textbf{Answer:} [NA] \\
\textbf{Justification:} The benchmark was annotated by professional
biomedical annotators rather than crowdsourced workers. The full
annotation guideline is included verbatim in
Appendix~\ref{app:guideline}.

\item \textbf{Institutional review board (IRB) approvals or
equivalent for research with human subjects.}
\emph{Does the paper describe potential risks incurred by study
participants, whether such risks were disclosed to the subjects, and
whether IRB approvals (or an equivalent approval/review based on the
requirements of your country or institution) were obtained?}\\
\textbf{Answer:} [NA] \\
\textbf{Justification:} The annotation work involves reading
publicly available scientific abstracts and producing structured
annotations; it does not involve human-subjects research as defined
under common-rule regulations. No IRB approval was required.

\item \textbf{Declaration of LLM usage.}
\emph{Does the paper describe the usage of LLMs if it is an
important, original, or non-standard component of the core methods
in this research?}\\
\textbf{Answer:} [Yes] \\
\textbf{Justification:} \texttt{Qwen2.5-7B-Instruct} is benchmarked
as the B3 baseline (\S\ref{sec:experimental_setup}). The paper makes
no use of LLMs in the construction of the benchmark itself:
tier-curation, deduplication, and gold-pool sampling are
deterministic and rule-based; the schema and annotation guideline
were authored by human contributors. A separate
\S\ref{sec:ai_use_statement} discloses the use of AI-assisted tools
during code drafting, paper editing, and LaTeX formatting; no AI
tool generated or modified the gold annotations, the experimental
results, or the scientific claims in this paper.

\end{enumerate}
